---
title: "Data Visualization"
subtitle: "Statistics for Political Science"
format: 
  clean-revealjs:
    echo: true
  html-math-method:
    method: mathjax
    url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
author:
  - name: Mason Auten
    email: mason.auten@vanderbilt.edu
    affiliations: Vanderbilt University
  - name: Patrick Buhr
    email: patrick.buhr@vanderbilt.edu
    affiliations: Vanderbilt University
date: today
self-contained: true
---

```{r}
#| echo: false
#| message: false
library("tidyverse")

## Only output five tibble rows by default
options(tibble.print_min = 5,
        tibble.max_extra_cols = 3)


#| echo: false
options(width = 42)
library(tidyverse)
```

```{r}
#| echo: FALSE
congress_data <- read_csv("congress_data.csv")

congress_data <- congress_data %>%
mutate(age = year - born,
gender = ifelse(female == 1, "Woman", "Man"),
party = ifelse(democrat == 1, "Democrat", "Republican"),
race = case_when(
latino == 1 ~ "Latino",
black == 1 ~ "Black",
aapi == 1 ~ "Asian",
TRUE ~ "White"
)) # fix this to include asian

untidy_election_data <- read_csv("untidy_election_data.csv")

long_data <- untidy_election_data %>%
select(!name) %>%
pivot_longer(cols = starts_with(c("presidentialvoteshare", "contributions")))

long_data <- separate(long_data,
col = "name",
into = c("var", "year"))

clean_election_data <- pivot_wider(long_data,
names_from = "var",
values_from = "value")
clean_election_data$year <- as.numeric(clean_election_data$year)
clean_election_data <- clean_election_data %>%
mutate(year = year + 1) %>%
rename(icpsr = ICPSR2)

congress_data <- left_join(congress_data, clean_election_data, by = c("icpsr", "year"))
congress_data <- congress_data %>%
mutate(presidentialvoteshare = ifelse(democrat==1, presidentialvoteshare, 100-presidentialvoteshare))
```

## Why Do We Visualize Data?

Why is it important that we visualize data?

Data visualization is most important at two stages:

1.  At the start of the research project, when we are trying to get a sense of our data.

2.  At the end of the research project, when we want to convey our findings to an audience.

Above all, think about your audience's cognitive process as they interpret your figure. 

## Why not just show a table?

If we were interested in the distribution of ideology (`dwnom1`) in Congress, why not just summarize?

::::: columns
::: {.column width="50%"}
```{r}
mean(congress_data$dwnom1, na.rm=TRUE)

sd(congress_data$dwnom1, na.rm=TRUE)

```
:::

::: {.column width="50%"}
```{r}
ggplot(congress_data, aes (x = dwnom1)) +
  geom_histogram(binwidth = 0.05)
```
:::
:::::

::: callout-tip
## General rule:
Your figure must convey more information to the audience than a table alone.
:::


## Histograms

### Continuous variables: Histogram

```{r}
#| echo: false

# Set default theme for plots
theme_set(theme_bw(base_size = 16))
```

```{r}
#| output-location: column
#| fig-width: 6.5
#| fig-height: 4.5
ggplot(congress_data, aes (x = dwnom1)) +
  geom_histogram(color = "black",
                 fill = "steelblue",
                 alpha = 0.5, # increases transparency
                 binwidth = 0.05) + # changes the number of bins
  labs(x = "First-Dimension DW-NOMINATE",
       y = "Number of Members of Congress",
       title = "Ideological Distribution of Members of Congress",
       subtitle = "Some are Liberal, Some are Conservative")
```

::: callout-warning
## ggplot syntax

ggplot commands are separated by addition `+`, not the pipe `%>%`.
:::

## Bar Charts

### Single Categorical Variable

```{r}
#| output-location: column
#| fig-width: 6.5
#| fig-height: 4.5
ggplot(congress_data, aes (y = gender)) +
  geom_bar(color = "black",
           fill = "steelblue",
           alpha = 0.7)+
  labs(x = "Number of Respondents",
       y = NULL,
       title = "Members of Congress by Gender")
```

## Bar Charts

### Two Categorical Variables

```{r}
#| output-location: column
#| fig-width: 6.5
#| fig-height: 4.5
ggplot(congress_data, aes(y = gender)) +
  geom_bar(aes(fill = party), position = "dodge") +
  labs(x = "Number of Respondents",
       y = NULL,
       fill = "Party",
       title = "Members of Congress by Gender and Party")
```

## Bar Charts

### Two Categorical Variables

```{r}
#| output-location: column
#| fig-width: 6.5
#| fig-height: 4.5
ggplot(congress_data, aes(y = gender)) +
  geom_bar(aes(fill = party), position = "dodge") +
  scale_fill_manual(values = c("Democrat" = "dodgerblue", "Republican" = "indianred")) +
  labs(x = "Number of Respondents",
       y = NULL,
       fill = "Party",
       title = "Members of Congress by Gender and Party")
```

::: callout-tip
## Changing colors or aesthetics

Use `scale_fill_manual()` to change aesthetics
:::

## Scatterplots

### Two Continuous Variables

```{r}
#| output-location: column
#| fig-width: 6.5
#| fig-height: 4.5
ggplot(congress_data,
       aes(x = dwnom1, y = presidentialvoteshare)) +
  geom_point() + # increases transparency
  labs( title = "Ideology and Presidential Vote Share",
        subtitle = "More Liberal Legislators Represent More Liberal Districts",
        x = "Ideology (Higher Values = More Conservative)",
        y = "Democratic Presidential Candidate \nVote Share in District"
# the \n tells ggplot to include a linebreak
)
```

## Scatterplots

### Two Continuous Variables

```{r}
#| output-location: column
#| fig-width: 6.5
#| fig-height: 4.5
ggplot(congress_data,
       aes(x = vote_share, y = presidentialvoteshare)) +
  geom_point(position = "jitter",
alpha = 0.5) + # increases transparency
  labs( title = "Ideology and Presidential Vote Share",
        subtitle = "More Liberal Legislators Represent More Liberal Districts",
        x = "Ideology (Higher Values = More Conservative)",
        y = "Democratic Presidential Candidate \nVote Share in District"
# the \n tells ggplot to include a linebreak
)+
  coord_cartesian(ylim = c(25, 100), xlim = c(45, 99))
```

::: callout-tip
## What if many observations take the same value?

`jitter()` creates minor perturbations if your data is bunching
:::

## Scatterplots

### Remove unopposed legislators

```{r}
#| output-location: column
#| fig-width: 6.5
#| fig-height: 4.5

congress_data %>%
  filter(vote_share<=95) %>%
ggplot(aes(x = vote_share, y = presidentialvoteshare)) +
  geom_point(position = "jitter",
alpha = 0.5) + # increases transparency
  labs( title = "Ideology and Presidential Vote Share",
        subtitle = "Liberal Legislators Represent Liberal Districts",
        x = "Ideology (Higher Values = More Conservative)",
        y = "Democratic Presidential Candidate \nVote Share in District"
# the \n tells ggplot to include a linebreak
)+
  coord_cartesian(ylim = c(25, 100), xlim = c(45, 99))
```

::: callout-tip
## Using `dplyr` with `ggplot`

You can also manipulate data before going into a ggplot, just remember to use `%>%` instead of `+`
:::

## Scatterplots

### facet

```{r}
#| output-location: column
#| fig-width: 6.5
#| fig-height: 4.5

congress_data %>%
  filter(vote_share<=95 & year %in% c(1981, 1985, 1989, 1993, 
                                      1996, 2001, 2005, 2009, 
                                      2013, 2017, 2021)) %>%
  mutate(election_year = year-1) %>%
ggplot(aes(x = vote_share, y = presidentialvoteshare)) +
  geom_point(position = "jitter",
alpha = 0.5) + # increases transparency
  labs( title = "Ideology and Presidential Vote Share",
        subtitle = "Elections Have Become More Correlated",
        x = "Ideology (Higher Values = More Conservative)",
        y = "Democratic Presidential Candidate \nVote Share in District"
# the \n tells ggplot to include a linebreak
)+
  facet_wrap(~ election_year,
ncol = 5)+
  coord_cartesian(ylim = c(25, 100), xlim = c(45, 99))
```

## Saving Plots

```{r}
nationalization_plot <- congress_data %>%
  filter(vote_share<=95 & year %in% c(1981, 1985, 1989, 1993, 
                                      1996, 2001, 2005, 2009, 
                                      2013, 2017, 2021)) %>%
  mutate(election_year = year-1) %>%
ggplot(aes(x = vote_share, y = presidentialvoteshare)) +
  geom_point(alpha = 0.5) + # increases transparency
  labs( title = "Ideology and Presidential Vote Share",
        subtitle = "Elections Have Become More Correlated",
        x = "Ideology (Higher Values = More Conservative)",
        y = "Democratic Presidential Candidate \nVote Share in District"
# the \n tells ggplot to include a linebreak
)+
  facet_wrap(~ election_year,
ncol = 3)+
  coord_cartesian(ylim = c(25, 100), xlim = c(45, 99))
```

## Nationalization Plot

### Testing

```{r}
#| echo: FALSE
nationalization_plot
```
