---
title: Directories and Coding
subtitle: "Statistics for Political Science"
format: clean-revealjs
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
author:
  - name: Mason Auten
    email: mason.auten@vanderbilt.edu
    affiliations: Vanderbilt University
  - name: Patrick Buhr
    email: patrick.smith@vanderbilt.edu
    affiliations: Vanderbilt University
date: today
self-contained: true
bibliography: skills_camp.bib
---

```{r}
#| echo: false
#| eval: true
#| output: false

pacman::p_load(tidyverse)
```


# Getting Started with R

## Work Space

When you open RStudio, you will see four different windows:

*  **Console**: where you will enter code, and where the outputs appear. You can type into it directly and press Enter to see the result.
*  **R Script**: Your script is where you keep a record of your work. As a general rule, you should always write your code here first before executing it in the console
*   **Environment and History:** The environment tab shows all of the objects that you currently have loaded into your workspace. The history tab shows a list of commands used so far.
*   **Files, plots, packages, help, viewer and presentation:** The files tab shows all the files and folders in your default workspace as if you were on a PC/Mac window. The plots tab will show all your graphs.


## Working Directories

The first thing to do is to set a working directory. Two broad ways to do this:


1. Manually
2. Project: Based

## Working Directories: Manually

- First, we want to check where the working directory is. There are two different commands that capture the same objective.

```{r}
#| echo: true
#| eval: true

getwd()

here::here()

```

* Next, we want to set that working directory 


```{r}
#| echo: true
#| eval: false

setwd("C:/Users/patri/OneDrive/Documents/MathCamp")
```

## Working Directories: Project Based

Working from a project allows you to set your working directory automatically.

::: {.incremental}

1. Click the project button
2. Select a new directory, then new project
3. Create the folder

:::

## Loading Packages

* The strength of R is in the community. 
* You'll frequently be loading packages that extend the baseline functions of R. 
* Installation only needs to be done once

```{r}
#| eval: false
#| echo: true
install.packages("tidyverse")

```

* However, we need to load packages before each use
  

```{r}
#| eval: false
#| echo: true
library(tidyverse)

```
- Every R script should start by loading your packages

## Executing Code

*  All scripts have active and inactive lines. Active lines are executable.
 *  On Mac, press COMMAND+RETURN
 *  On PC, press CTRL+RETURN


```{r}
#| eval: true
#| echo: true
2+2

```

* While R is flexible and powerful, it can also do simple calculations

```{r}
#| eval: true
#| echo: true
2+2
6-2
250*5

```

## Executing Code Cont. 
* Commenting out code will cause it to be skipped

```{r}
#| eval: true
#| echo: true


## Commenting out lines means that R won't try to run that line
## Notice how R skips over this line? 
3^2
```

* Comment. Your. Code. 
 * It saves future you and future reviewers headaches

## Objects and Functions 

* R is a vector-based language. 
* We can store values as "objects"

```{r}
#| eval: false
#| echo: true

NAME <- VALUE
```

* Any of the previous calculations could be stored as objects

```{r}
#| eval: true
#| echo: true

add_obj <- 4+4 
mult_obj <- 250*5
```

* Names of objects are case sensitive
 * Where a space would go, try using `.` or ` _` instead. 

## Objects and Functions Cont. {.scrollable}

* To print the values of the objects, we simply list them 

```{r}
#| eval: true
#| echo: true

add_obj 
mult_obj 
```

* For large vectors, we can also specify how many of the results are printed

```{r}
#| eval: true
#| echo: true

# With large objects, it helps to tell it how many values you want printed

mult_obj  %>% 
head(10)
```

## Objects and Functions Cont. {.scrollable}
* To see all objects in your working environment: 

```{r}
#| eval: true
#| echo: true

ls()

```

* To remove a specific object 

```{r}
#| eval: true
#| echo: true
rm(add_obj)

# Or remove all objects with rm(list=ls())
```

* Objects in R are separated into "classes"
 * Classes include: character, numeric, vector, factor, matrix, list, dataframe, and so on. 


```{r}
#| eval: true
#| echo: true

class(mult_obj)
```

## Objects and Functions Cont. {.scrollable}

* Again, R is a vector-based language
 * We work with an $n \times 1$ collection of elements
* Here is one way to create vectors

```{r}
#| eval: true
#| echo: true

ranking <- c(1,2,3,4,5)
ranking
class(ranking)

```

- We can create character-based objects and add information to it ex post


```{r}
#| eval: true
#| echo: true

our_school <- "Vanderbilt"

class(our_school)

sec_schools <- c(our_school, "University of Florida", "UT Austin", "Georgia", "Texas A&M")
sec_schools
class(sec_schools)

```

## Objects and Functions Cont. {.scrollable}
- We can then check the length of this particular vector 

```{r}
#| eval: true
#| echo: true
length(ranking)

length(sec_schools)
```

## Objects and Functions Cont. {.scrollable}
- Each element of a vector must be of the same class

```{r}
#| eval: true
#| echo: true


class(ranking)

class(sec_schools)

mixed_vector <- c(ranking, sec_schools)

mixed_vector

class(mixed_vector)
```

## Data Frames{.scrollable}

- Most functions we use in R utilize dataframes
- Dataframes do not have to be of the same class
 - But each element in a column should be of the same column 

```{r}
#| eval: true
#| echo: true


sec_rankings <- cbind.data.frame(ranking, sec_schools)

class(sec_rankings)

sec_rankings

```

## Data Frames Cont. {.scrollable}
- The dataframe is treated as a type of matrix, and we can access specific elements like so 


```{r}
#| eval: true
#| echo: true

sec_rankings[2,2]

# Another way to do it
sec_rankings[2,"sec_schools"]
```

- Access a specific column like so 


```{r}
#| eval: true
#| echo: true


sec_rankings$sec_schools
```

## Functions 

- R functions are analogous to mathematical functions. They are a mapping from one set of values to another
- Functions are made of three components [@wickham2019Advanced] 
 1. The arguments you use
 2. The body of code inside the function
 3. The structure of the data 
- Calling a function `function(input)`


```{r}
#| eval: false
#| echo: true


?function()

```


## Mathematical Operators 

- Let's go over some basic mathematical operators you'll use

```{r}
#| eval: true
#| echo: true

# Square Root
sqrt(144)


# Natural Log
log(10)

# Exponent
exp(3)

```

## Mathematical Operators Cont. 

```{r}
#| eval: true
#| echo: true

# sequence of numbers
seq(from = 1, to = 5, by = 0.5)

# R also has a shortcut for sequences of integers
1:5

# Repeating the same input many times
rep(x = c(1, 2, 3), times = 3)


```

* We can store data as an object and use it in operations as well 

```{r}
#| eval: true
#| echo: true
numbers <- 1:5
sum(numbers)
```

## Logical Statements

- To write more advanced code, you'll want to get familiar with the way that R handles logical statements. 
- Logical statements are used in settings where you want to use a TRUE/FALSE logic. 
- `==` means "is exactly true." 

```{r}
#| eval: true
#| echo: true

2 == 2

2 == 3

2 == "Joe Biden"

```

## Logical Statements Cont. 
- To say that a particular object is not equal to another, we use `!=`

```{r}
#| eval: true
#| echo: true

# it is FALSE that 2 is not equal to 2
2 != 2

# it is TRUE that 2 is not equal to 3
2 != 3
```

## Logical Statements Cont. 
- We can also use strict inequalities and weak inequalities 
- Here are some examples of strict inequalities: 

```{r}
#| eval: true
#| echo: true

# 3 is greater than 2
3 > 2

# 3 is not less than 2
3 < 2

# 2 is less than 3
2 < 3

```

## Logical Statements Cont.
```{r}
#| eval: true
#| echo: true

# 3 is greater than 2
3 > 2

# 3 is not less than 2
3 < 2

# 2 is less than 3
2 < 3

```


## Logical Statements Cont. 

- We can use weak inequalities by adding a `=` in front of or behind `>`

```{r}
#| eval: true
#| echo: true


# it is FALSE that 3 is greater than 3
3 > 3

# but it is TRUE that 3 is greater than or equal to 3
3 >= 3
```

- Adding a `!` before your logical statement creates a negation, turning TRUE into FALSE and vice versa. 

```{r}
#| eval: true
#| echo: true

# it is FALSE that 2 is greater than 3
2 > 3

# but it is TRUE that 2 is NOT greater than 3
!(2 > 3)


```

## Logical Statements Cont. 

- We can combine these operators to create compound logical statements. 
- R returns true if and only if every condition is true
- Specify multiple conditions using the `&` symbol.

## Logical Statements Cont. 
- Example:
```{r}
#| eval: true
#| echo: true

# creating an object called "number" that takes the value of 5
number <- 5

# it is TRUE that number is not equal to 3
number != 3

# it is FALSE that number is greater than 7
number > 7

# it is FALSE that number is not equal to 3 AND that number is greater than 7
number !=3 & number > 7

# however, it is TRUE that number is not equal to 3 and
# that number is greater than 4
number !=3 & number > 4
```

## Logical Statements Cont. 
- `|` means "OR" in R. Use it to return true if one of the conditions specified in the or statement is true. 

```{r}
#| eval: true
#| echo: true
# it is TRUE that 5 is not equal to 3
5 != 3

# it is FALSE that 5 is greater than 7
5 > 7

# even though 5 is not greater than 7, 5 is not equal to 3. 
# therefore, it is 
```




## References

::: {#refs}
:::