---
title: Introduction Continued 
subtitle: "Introduction to Graduate Research"
format: clean-revealjs
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
author:
  - name: Mason Auten
    email: mason.auten@vanderbilt.edu
    affiliations: Vanderbilt University
  - name: Patrick Buhr
    email: patrick.buhr@Vanderbilt.Edu
    affiliations: Vanderbilt University
date: today
self-contained: true
bibliography: skills_camp.bib
---

```{r}
#| echo: false
#| eval: true
#| output: false

pacman::p_load(tidyverse)
```

# Programming

## Why Use R for Statistics?

-   R is a powerful and widely used programming language for statistical analysis.

-   It is free, open-source, and widely supported by researchers and data analysts.

-   R provides extensive libraries for data visualization, regression modeling, and machine learning.

-   Learning R enhances your ability to conduct independent research and analyze real-world data.

## Principles of Programming

**Principle One:** "Write code for humans."

-   Indent your code according to style conventions in www.style.tidyverse.org. RStudio will typically do this automatically.\
-   Add explanatory comments to everything you do using `#`.\
-   Basic principle: if I got hit by a bus tomorrow, could people figure out what I was trying to do and finish my code?

## Principles of Programming

**Principle Two:** "Let the computer do the work"

-   Copy-and-paste doesn't work: prone to errors, time intensive if you need to make changes, and confusing to read.\
-   Any repetitive task can be done with minimal lines of code.\
-   If someone else has your data and code, they must be able to replicate your same results.

## Debugging and Asking for Help

::::: columns
::: {.column width="50%"}
Types of problems:

-   Errors: the R script stops because it cannot execute your command.\
-   Warnings: the R script was able to execute your command (and produce output) but thinks you need to check the output.\
-   Unexpected output: your script executed without warnings, but the output isn't what you expected (e.g. output is NA).
:::

::: {.column width="50%"}
Solving problems:

1.  Find the exact line of code causing the problem.\
2.  Look on StackOverflow/ChatGPT.\
3.  Come to us in office hours.
:::
:::::

## Statement on ChatGPT

-   ChatGPT and other LLMs are **tools**. Being able to use them well will be an essential part of your workflow in the future.
-   However, like any tool, it is exceptionally easy to misuse ChatGPT if you don't know what you are doing.
-   Use ChatGPT for learning, to clarify concepts, and to troubleshoot your own code—but do **not** use it to replace your own analytical thinking.
-   Your ability to think clearly and rigorously about the underlying statistics is what will set you apart.

# Getting Started with R

## Work Space

When you open RStudio, you will see four different windows:

-   **Console**: where you will enter code, and where the outputs appear. You can type into it directly and press Enter to see the result.
-   **R Script**: Your script is where you keep a record of your work. As a general rule, you should always write your code here first before executing it in the console
-   **Environment and History:** The environment tab shows all of the objects that you currently have loaded into your workspace. The history tab shows a list of commands used so far.
-   **Files, plots, packages, help, viewer and presentation:** The files tab shows all the files and folders in your default workspace as if you were on a PC/Mac window. The plots tab will show all your graphs.

## Working Directories

The first thing to do is to set a working directory. Two broad ways to do this:

1.  Manually
2.  Project: Based

## Working Directories: Manually

-   First, we want to check where the working directory is. There are two different commands that capture the same objective.

```{r}
#| echo: true
#| eval: true

getwd()

here::here()

```

-   Next, we want to set that working directory

```{r}
#| echo: true
#| eval: false

setwd("C:/Users/patri/OneDrive/Documents/MathCamp")
```

## Working Directories: Project Based

Working from a project allows you to set your working directory automatically.

::: incremental
1.  Click the project button
2.  Select a new directory, then new project
3.  Create the folder
:::

## Loading Packages

-   The strength of R is in the community.
-   You'll frequently be loading packages that extend the baseline functions of R.
-   Installation only needs to be done once

```{r}
#| eval: false
#| echo: true
install.packages("tidyverse")

```

-   However, we need to load packages before each use

```{r}
#| eval: false
#| echo: true
library(tidyverse)

```

-   Every R script should start by loading your packages

## Executing Code

-   All scripts have active and inactive lines. Active lines are executable.
-   On Mac, press COMMAND+RETURN
-   On PC, press CTRL+RETURN

```{r}
#| eval: true
#| echo: true
2+2

```

-   While R is flexible and powerful, it can also do simple calculations

```{r}
#| eval: true
#| echo: true
2+2
6-2
250*5

```

## Executing Code Cont.

-   Commenting out code will cause it to be skipped

```{r}
#| eval: true
#| echo: true


## Commenting out lines means that R won't try to run that line
## Notice how R skips over this line? 
3^2
```

-   Comment. Your. Code.
-   It saves future you and future reviewers headaches

## Objects and Functions

-   R is a vector-based language.
-   We can store values as "objects"

```{r}
#| eval: false
#| echo: true

NAME <- VALUE
```

-   Any of the previous calculations could be stored as objects

```{r}
#| eval: true
#| echo: true

add_obj <- 4+4 
mult_obj <- 250*5
```

-   Names of objects are case sensitive
-   Where a space would go, try using `.` or `_` instead.

## Objects and Functions Cont. {.scrollable}

-   To print the values of the objects, we simply list them

```{r}
#| eval: true
#| echo: true

add_obj 
mult_obj 
```

-   For large vectors, we can also specify how many of the results are printed

```{r}
#| eval: true
#| echo: true

# With large objects, it helps to tell it how many values you want printed

mult_obj  %>% 
head(10)
```

## Objects and Functions Cont. {.scrollable}

-   To see all objects in your working environment:

```{r}
#| eval: true
#| echo: true

ls()

```

-   To remove a specific object

```{r}
#| eval: true
#| echo: true
rm(add_obj)

# Or remove all objects with rm(list=ls())
```

-   Objects in R are separated into "classes"
-   Classes include: character, numeric, vector, factor, matrix, list, dataframe, and so on.

```{r}
#| eval: true
#| echo: true

class(mult_obj)
```

## Objects and Functions Cont. {.scrollable}

-   Again, R is a vector-based language
-   We work with an $n \times 1$ collection of elements
-   Here is one way to create vectors

```{r}
#| eval: true
#| echo: true

ranking <- c(1,2,3,4,5)
ranking
class(ranking)

```

-   We can create character-based objects and add information to it ex post

```{r}
#| eval: true
#| echo: true

our_school <- "Vanderbilt"

class(our_school)

sec_schools <- c(our_school, "University of Florida", "UT Austin", "Georgia", "Texas A&M")
sec_schools
class(sec_schools)

```

## Objects and Functions Cont. {.scrollable}

-   We can then check the length of this particular vector

```{r}
#| eval: true
#| echo: true
length(ranking)

length(sec_schools)
```

## Objects and Functions Cont. {.scrollable}

-   Each element of a vector must be of the same class

```{r}
#| eval: true
#| echo: true


class(ranking)

class(sec_schools)

mixed_vector <- c(ranking, sec_schools)

mixed_vector

class(mixed_vector)
```

## Data Frames {.scrollable}

-   Most functions we use in R utilize dataframes
-   Dataframes do not have to be of the same class
-   But each element in a column should be of the same column

```{r}
#| eval: true
#| echo: true


sec_rankings <- cbind.data.frame(ranking, sec_schools)

class(sec_rankings)

sec_rankings

```

## Data Frames Cont. {.scrollable}

-   The dataframe is treated as a type of matrix, and we can access specific elements like so

```{r}
#| eval: true
#| echo: true

sec_rankings[2,2]

# Another way to do it
sec_rankings[2,"sec_schools"]
```

-   Access a specific column like so

```{r}
#| eval: true
#| echo: true


sec_rankings$sec_schools
```

## Functions

-   R functions are analogous to mathematical functions. They are a mapping from one set of values to another
-   Functions are made of three components [@wickham2019Advanced]

1.  The arguments you use
2.  The body of code inside the function
3.  The structure of the data

-   Calling a function `function(input)`

```{r}
#| eval: false
#| echo: true


?function()

```

## Mathematical Operators

-   Let's go over some basic mathematical operators you'll use

```{r}
#| eval: true
#| echo: true

# Square Root
sqrt(144)


# Natural Log
log(10)

# Exponent
exp(3)

```

## Mathematical Operators Cont.

```{r}
#| eval: true
#| echo: true

# sequence of numbers
seq(from = 1, to = 5, by = 0.5)

# R also has a shortcut for sequences of integers
1:5

# Repeating the same input many times
rep(x = c(1, 2, 3), times = 3)


```

-   We can store data as an object and use it in operations as well

```{r}
#| eval: true
#| echo: true
numbers <- 1:5
sum(numbers)
```

## Logical Statements

-   To write more advanced code, you'll want to get familiar with the way that R handles logical statements.
-   Logical statements are used in settings where you want to use a TRUE/FALSE logic.
-   `==` means "is exactly true."

```{r}
#| eval: true
#| echo: true

2 == 2

2 == 3

2 == "Joe Biden"

```

## Logical Statements Cont.

-   To say that a particular object is not equal to another, we use `!=`

```{r}
#| eval: true
#| echo: true

# it is FALSE that 2 is not equal to 2
2 != 2

# it is TRUE that 2 is not equal to 3
2 != 3
```

## Logical Statements Cont.

-   We can also use strict inequalities and weak inequalities
-   Here are some examples of strict inequalities:

```{r}
#| eval: true
#| echo: true

# 3 is greater than 2
3 > 2

# 3 is not less than 2
3 < 2

# 2 is less than 3
2 < 3

```

## Logical Statements Cont.

```{r}
#| eval: true
#| echo: true

# 3 is greater than 2
3 > 2

# 3 is not less than 2
3 < 2

# 2 is less than 3
2 < 3

```

## Logical Statements Cont.

-   We can use weak inequalities by adding a `=` in front of or behind `>`

```{r}
#| eval: true
#| echo: true


# it is FALSE that 3 is greater than 3
3 > 3

# but it is TRUE that 3 is greater than or equal to 3
3 >= 3
```

-   Adding a `!` before your logical statement creates a negation, turning TRUE into FALSE and vice versa.

```{r}
#| eval: true
#| echo: true

# it is FALSE that 2 is greater than 3
2 > 3

# but it is TRUE that 2 is NOT greater than 3
!(2 > 3)


```

## Logical Statements Cont.

-   We can combine these operators to create compound logical statements.
-   R returns true if and only if every condition is true
-   Specify multiple conditions using the `&` symbol.

## Logical Statements Cont.

-   Example:

```{r}
#| eval: true
#| echo: true

# creating an object called "number" that takes the value of 5
number <- 5

# it is TRUE that number is not equal to 3
number != 3

# it is FALSE that number is greater than 7
number > 7

# it is FALSE that number is not equal to 3 AND that number is greater than 7
number !=3 & number > 7

# however, it is TRUE that number is not equal to 3 and
# that number is greater than 4
number !=3 & number > 4
```

## Logical Statements Cont.

-   `|` means "OR" in R. Use it to return true if one of the conditions specified in the or statement is true.

```{r}
#| eval: true
#| echo: true
# it is TRUE that 5 is not equal to 3
5 != 3

# it is FALSE that 5 is greater than 7
5 > 7

# even though 5 is not greater than 7, 5 is not equal to 3. 
# therefore, it is 
```

# Load Data

## Reading in Data

-   Most of the data you work with will be in CSV files

```{r}
#| echo: TRUE

library(tidyverse)

congress_data <- read_csv(here::here("inst", "congress_data.csv"))

```

## Reading in Stata Files

-   You can also read in files made in Stata using the packge `haven`

```{r}
#| eval: false
#| echo: true
library(haven)
congress_data <- read_dta("congress_data.dta")
```

-   Or you can read in data from a URL.

```{r}
#| eval: false
#| echo: true

congress_data <- read_dta("www.thelawmakers.org/uploads/congress_data.dta")
```

## Variables {.scrollable}

-   We can list the name of all the variables using `ls()`
-   We can extract a specific variable by using `$`

```{r}
#| echo: true
ls(congress_data)

congress_data$seniority

head(congress_data$seniority, 20)

```

## Variables Cont.

-   Use square brackets `[]` to extract individual values:

-   And use `:` within the brackets to extract a range:

```{r}
#| echo: true
congress_data$seniority[5]

congress_data$seniority[1:10]

```

## References

::: {#refs}
:::