---
title: "Introduction"
subtitle: "Introduction to Graduate Research"
format: 
  clean-revealjs:
    echo: true
  html-math-method:
    method: mathjax
    url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
author:
  - name: Mason Auten
    email: mason.auten@vanderbilt.edu
    affiliations: Vanderbilt University
  - name: Patrick Buhr
    email: patrick.buhr@vanderbilt.edu
    affiliations: Vanderbilt University
date: today
self-contained: true
---

## Loading in Data
```{r}
#| echo: false
#| message: false
library("tidyverse")

## Only output five tibble rows by default
options(tibble.print_min = 5,
        tibble.max_extra_cols = 3)
```
:::::: columns
::: {.column width="50%"}
1.  Navigate to your working directory.
2.  Download the `congress_data.csv` file and put it there
3.  Set as R's working directory
    -   Windows: `setwd('C:/path/to/directory')`
    -   Mac: `setwd('~/path/to/directory')`
    -   ... or just navigate in RStudio and set it that way
4.  Check that `congress_data <- read_csv("congress_data.csv")` works
:::

::: {.column width="5%"}
:::

::: {.column width="45%"}
```{r}
#| echo: false
options(width = 42)
library(tidyverse)
```

```{r}
#| echo: TRUE
congress_data <- read_csv("congress_data.csv")

print(congress_data)
```
:::
::::::

# Manipulating data

## Basics of data manipulation

Use `$` to extract a single column

```{r}
#| echo: false
options(width = 93, max.print = 25)
```

```{r}
congress_data$name
```

Use square brackets `[` `]` to extract individual value(s)

```{r}
congress_data$name[5]     # age of the 5'th row of the data
congress_data$name[1:10]  # first 10 ages in the data
```

## Describing Data

On Tuesday, we learned a variety of options for descriptive statistics.

```{r}
mean(congress_data$seniority, na.rm = TRUE)    # average/mean
median(congress_data$seniority, na.rm = TRUE)  # median
sd(congress_data$seniority, na.rm = TRUE)      # standard deviation
cor(congress_data$seniority, congress_data$les, use = "complete.obs")
```

## Manipulating Data

In your career as a social scientist, you will spend much more time cleaning, merging, and tidying data than actually running analysis.

Fortunately, the `dplyr` package within `tidyverse` gives many options for data manipulation.

We will run through the most popular here.

## `rename()`

`rename()` allows you to rename variables to different names.

Suppose we think that the variable "female" is a bit dated.

```{r}
congress_data %>% rename(woman = female)
```

## `filter()`

You can also filter out rows that meat or do not meat a certain criteria.

Maybe we are only interested in Members of Congress who are Millenials (born after 1981)

```{r}
congress_data %>% filter(born <= 1981)
```

## `select()`

You can also subset to certain columns using the `select()` command.

Let's only keep a Member of Congress's name, birthyear, and first-dimension DW-NOMINATE score (a commonly used measure of ideology.)

```{r}
congress_data %>% select(name, born, dwnom1)
```

## The Pipe `%>%`

We can chain together commands using the `%>%` function.

For example, we can write the above code as:

```{r}
congress_data %>% filter(born >= 1981) %>%
  select(name, born, dwnom1)
```

::: callout-tip
## Different pipes

You may also see the pipe written as `|>` for R's build-in function. Both are essentially equivalent.
:::

## `mutate()`

We can change or create new variables using the `mutate()` function. 

We can create new variables that are transformations of existing variables.

For example, we can create a variable for a legislator's age as the difference between their birth year and the current year:
```{r}
congress_data <- congress_data %>%
  mutate(age = year - born)

congress_data %>% select(born, year, age)
```

## `mutate()`

We can change or create new variables using the `mutate()` function. 

We can create new variables using an `ifelse()` statement. 

For example, maybe we want to create a new categorical variable for gender. 
```{r}
congress_data <- congress_data %>%
  mutate(gender_binary = if_else(female == 1, "Female", "Male"))

congress_data %>% select(female, gender_binary)
```

## `mutate()`

If you have more than one category, `ifelse()` can become unwieldy. This is a great use for the `case_when()` function.

For example, maybe we want to create a new categorical variable for race. 
```{r}
congress_data <- congress_data %>%
  mutate(race = case_when(
black == 1 ~ "Black",
latino == 1 ~ "Latino",
aapi == 1 ~ "Asian",
TRUE ~ "White"
))
```

:::{.callout-warning}
## Dangers of `ifelse()` and `case_when()`
Who else gets coded as "White" here?
:::

## `mutate()`

We can mutate several variables together in the same function by placing a comma after each line:

```{r}
congress_data <- congress_data %>%
  mutate(age = year - born,
         gender_binary = if_else(female == 1, "Female", "Male"),
         race = case_when(
           black == 1 ~ "Black",
           latino == 1 ~ "Latino",
           aapi == 1 ~ "Asian",
           TRUE ~ "White"
           ))

congress_data %>% select(age, gender_binary, race)
```

## Summaries by group

You can create summaries of different variables using the `summarize()` function.
```{r}
congress_data %>% summarize(average_effectiveness = mean(les, na.rm=T))
```

Summaries are most useful when combined with `group_by()` which divides the sample into different groups and then performs subgroup analysis.
```{r}
congress_data %>%
  group_by(majority) %>%
  summarize(number_of_respondents = n(),
            average_effectiveness = mean(les, na.rm=T))
```

## Summaries by group

We can also create summaries with multiple groups.

```{r}
congress_data %>%
  group_by(majority, female) %>%
  summarize(number_of_respondents = n(),
            average_effectiveness = mean(les, na.rm=T))
```


:::{.callout-warning}
## Summaries
If you overwrite your original dataset, `summarize()` will reduce the dimensions of your dataset to only what you include.  
:::

# 10 Minute Break

## Tidying Data

- Let's go over how to tidy data that isn't immediately ready to use. 


```{r}
untidy_election_data <- read_csv("untidy_election_data.csv")

head(untidy_election_data)
```

- Each row or observation of `untidy_election_data` is a Member of Congress. 

## `pivot()`

- We can take data and pivot it either wider or longer 


::: {.callout-important}
- In wide data, each row is an observation and each variable is a column. Observations should not be repeated.  
- In long data, values often repeat across rows, and data is stored in a more compact, stacked format. One column identifies the unit (e.g., `state`), another specifies the variable name (e.g., `variable`), and a third holds the corresponding value. 
:::

