# Getting Started

R is a programming language widely used in statistical analysis. Learning R can come with a steep learning curve, but once you get the hang of it there's a nearly endless number of exciting things you can use it for. Time invested in deepening your R skills will never be time wasted.

Code in R, and any other programming language, is a series of instructions to the computer. Usually, you will begin by either loading or creating objects, for example by importing a dataset you downloaded or defining a new variable. You'll then do various things to or with those objects, for example removing a row from a matrix, running a regression using your dataset, or making a graph. This is done using functions; a huge number of functions are built into R, many more are available through thousands of packages, and you will also write your own functions.

## The R Workspace

When you open RStudio, you will see four different windows:

-   **Console**: where you will enter code, and where the outputs appear. You can type into it directly and press enter to see the result.

-   **R Script**: your script is where you keep a record of your work. As a general rule, you should always write your code here first before executing it in the console. Try to write your code as clean and well-structured as possible. You want yourself and other scholars to understand your code and be able to replicate it at a different point in time.

-   **Environment and History:** The environment tab shows all of the objects that you currently have loaded into your workspace. The history tab shows a list of commands used so far.

-   **Files, plots, packages, help, viewer and presentation:** The files tab shows all the files and folders in your default workspace as if you were on a PC/Mac window. The plots tab will show all your graphs. The packages tab will list a series of packages or add-ons needed to run certain processes. It shows you which packages you have currently installed and loaded. For additional info on a specific command, refer to the help tab.

If you don't have a file set to open when you first open RStudio, you will likely only see three of these windows when RStudio first launches (Console, Environment and History, and Files, plots, packages, help, and viewer). Once you open a code file, the R Script window will also appear.

## Setting a Working Directory
The first thing you'll want to do when working in R is to set a working directory. This is the folder on your computer that R will default to for saving and loading files (unless told otherwise). In other words, this is where R will look for files you call and is where R will save files you create. You can check where your working directory currently is using `getwd()`

```{r} 
#| echo: true

getwd()
```

To change this, we use the `setwd()` function. Below is the line of code I am using to set my working directory. Yours will be unique to your file organization. If you don't want to memorize the file path to set your working directory, you can search for one. To do this, select Session from the top menu, then Set Working Directory, then Choose Directory. Now you can choose which folder you would like to use as your working directory, which R will update automatically. It is a good idea to copy and paste the command that sets your working directory directly into your R script.

```{r}
#| eval: false
#| echo: true
setwd("C:/Users/patri/OneDrive/Documents/MathCamp")}
```

You'll notice that the `setwd()` command contains a specific file path on my personal laptop. This is not ideal for a script that I want a collaborator to be able to use. If the folder is in a shared directory like Dropbox, you can write a more universal command which would work for anyone who had access to this folder.

## Loading Packages

R has a number of functions built in, but you will often find yourself needing a function that is not included in base R. You access these functions by downloading and loading packages. You only need to download packages once. To install a package, use the `install.packages()` command, with the package name in quotation marks. For most of our work today, and most of your work in general, we'll be using the `tidyverse` suite of packages. `tidyverse` is convenient because it allows you to download many commonly used packages at once.

```{r}
#| eval: false 
#| echo: true
install.packages("tidyverse")
```

You only need to install packages once, but will need to load packages from your library every time you use them. Note that you do not need to use quotations when loading packages.

```{r}
#| eval: false 
#| echo: true
library(tidyverse)
```

Start every script by first setting your working directory and then loading in all packages that you are going to use.

## Executing Code

Scripts of code have active and inactive lines. Active lines are executable. To execute a section of code, select the entire section, and then press COMMAND+RETURN on a Mac, or CTRL+RETURN on a PC. Pressing these keys without selecting something will execute the line where your cursor is. Alternatively, you can select the portion of code you want to execute and hit "Run" in the top right corner of the code window. You can also execute all of the code in a .R file by pressing the "Source" button.

```{r}
2+2

```

Notice the features of the output R produces. First we see `[1]`. This indicates that whatever follows is the first element of the output. Sometimes output will have multiple elements, but here there is just one. Then we see the contents of that output: the number 4. If something has gone wrong in executing the code, R will print an error message instead of the output, which you can use to figure out what the problem is.

We can perform other arithmetic operations using R, such as subtraction...

```{r}
6-2
```

...division...

```{r}
60/10
```

...or multiplication

```{r}
250*5
```

Inactive or "commented" lines of code begin with a hashtag (or pound) mark, #, and R will skip over them if you tell it to execute your script. Commenting using the `#` mark is useful for annotating a script file and making notes with collaborators. Once you type a `#` symbol, R will comment out (disregard) all subsequent characters until you hit return and start a new line.

```{r}
# <- notice that I have included a hashtag
# R will ignore any text with a proceeding hashtags
# commenting allows you to explain code and guide the reader
# my next line of code calculates 3 squared
3^2
```

Comments are crucial to writing good code. Keeping clean and annotated script files is important for many reasons. It helps avoid errors, it ensures your work is replicable, and it enables both your colleagues and your future self to understand what the code is doing. Get into the habit of always providing comments explaining your code.
