---
title: "Problem Set 3 Answers"
author: "YOUR NAME GOES HERE"
date: 2025-03-19
format:
  pdf:
    fontsize: 12pt
    highlight-style: tango
---


# Part A: Instrumental variables concepts

## 1. Key components of the analysis

Answer goes here.

## 2. Presence of selection bias

Answer goes here.

## 3. Instrument and its validity

Answer goes here.

## 4. Compliers

Answer goes here.

## 5. Conclusions and evidence

Answer goes here.

# Part B: Instrumental variables in practice

```{r}
#| message: false
options(width = 72)
library("tidyverse")
library("broom")
library("AER")
df_dlo <- read_csv("data_set4_dlo.csv")
```

## 1. Balance across treatment

```{r}
# Code goes here
```

Interpretation goes here.

## 2. Balance across instrument

```{r}
# Code goes here
```

Interpretation goes here.

## 3. First stage and reduced form

```{r}
# Code goes here
```

Interpretation goes here.

## 4. IV without controls

```{r}
# Code goes here
```

Interpretation goes here.

## 5. IV with De La O's controls

```{r}
# Code goes here
```

Interpretation goes here.

## 6. IV with your own controls

```{r}
# Code goes here
```

Interpretation goes here.

